---
title: "Automated Report Tables"
author: "Statistics Student"
format:
  html:
    self-contained: true
    self-contained-math: true
params:
  species: "Adelie"
---



```{r}
#| echo: false
#| message: false
#| warning: false
# this block loads R packages that may be needed for the analysis.
library(tidymodels)
library(tidyverse)
library(gtsummary)
library(palmerpenguins)
```



## Analysis Report

This report describes the distribution of body mass for penguin species `r params$species`.

```{r}
#| echo: false
penguins_filter <- penguins %>%
  dplyr::filter(species == params$species, !is.na(body_mass_g)) 
diff_rows <- nrow(penguins) - nrow(penguins_filter)
```

We have data on `r nrow(penguins)` penguins. Only
`r nrow(penguins_filter)` are classified as species `r params$species`. 
The distribution of these `r params$species` penguins 
are shown below:

```{r}
#| echo: false
ggplot(data=penguins_filter, mapping=aes(x=body_mass_g)) +
  geom_histogram(binwidth=60) +
  labs(title = paste(params$species, "population"))
```

```{r}
#| label: tbl-gtsumm
#| tbl-cap: "Summary Stats Table"
#| echo: false
#| message: false
#| warning: false
penguins_filter %>%
  dplyr::select(body_mass_g) %>%
  tbl_summary(data=., 
              label = list(body_mass_g ~ "Body Mass (grams)"),
              statistic = list(all_continuous() ~ "{median} ({p25},{p75})"),
              type = list(body_mass_g ~ "continuous")) %>%
  modify_header(label ~ "**Variable**") %>%
  bold_labels()

```


Summary statistics for distribution of the `r params$species` penguins 
is shown in  @tbl-gtsumm .
